---
title: "Replicación de Global inequality remotely sensed"
author: "Ignacio Sarmiento-Barbieri"
date: "2024-06-12"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
source("00_packages.R")
knitr::opts_chunk$set(echo = TRUE)
here::here()
```

## Introduction

Over the past decades, there has been a transition from theoretical to data-driven inequality research (1). However, progress is limited by a lack of data on economic prosperity at the household level (2) as well as the absence of consensus on ways of measuring economic inequality (3, 4). Practical constraints include limited coverage, incomparability at population subscales, dependence on misreported income surveys, and low-quality data collection in developing economies (5–7). 

# Cargamos los datos
```{r}
load(here("../02_analysis/01_input",'pnas_light_ineq.RData'))
```


# Replicar Tabla 1

```{r echo=FALSE, warning=FALSE,message=FALSE, results='hide'}
mod_1 <- lmer(Gini ~ Light_Gini + 
                (1 | Year), 
              data = df_nat)
summary(mod_1)
r.squaredGLMM(mod_1)[2]
AIC(mod_1)
AICc(mod_1)

mod_2 <- lmer(Gini ~ Light_Gini +  
                log(POP) +
                (1 | Year), 
              data = df_nat)
summary(mod_2)
r.squaredGLMM(mod_2)[2]
AIC(mod_2)
AICc(mod_2)


mod_3 <- lmer(Gini ~ Light_Gini +  
                  log(POP) +
                  log(GDP) + 
                  (1 | Year), 
                data = df_nat)
summary(mod_3)
r.squaredGLMM(mod_3)[2]
AIC(mod_3)
AICc(mod_3)

mod_4 <- lm(Gini ~ Light_Gini +  
                log(POP) +
                log(GDP), 
              data = df_nat)
summary(mod_4)
```

Esto sí es Markdown, porque podemos poner **negrilla**, *cursiva*, y ~~tachar~~. 

$$
\ln (wage) = \alpha + \beta_1 age + \beta_2 age^2
$$


```{r, results='asis', echo=TRUE, warning=FALSE}
modelsummary(list(mod_1,mod_2,mod_3,mod_4), output="latex")
```

```{r, echo=F}
# 
# #"html"
# #"markdown
# #modelsummary(models, output = "table.docx")
# 
# modelsummary(list(mod_1,mod_2,mod_3,mod_4), fmt=2, "html")
# modelsummary(list(mod_1,mod_2,mod_3,mod_4), 
#              fmt=2,
#              estimate  = "{estimate}{stars} ({std.error})",
#              statistic=NULL,
#              stars = c('***' = .001, '**' = .05,"*" =0.1), 
#              "html")
```


# Replicar Figura 1
```{r fig1prep, echo=FALSE}
mod_3 <- lmer(Gini ~ Light_Gini +  
                log(POP) +
                log(GDP) + 
                (1 | Year), 
              data = df_nat)
```


```{r fig1disp1, }

#Fig 1
visreg(mod_3, 'Light_Gini', gg = T, overlay = T, band = F) + 
  geom_point(size = 2, colour = 'grey') + labs(x = 'Light Gini', y = 'Income Gini')

#los ordenes estan invertidos
```

```{r fig1disp2, echo=FALSE,  out.width="60%", fig.cap="A nice image."}
spplot(world_2010, c("gini_disp"), col.regions = colorRampPalette(brewer.pal(9, 'YlOrRd'))(20), 
       cuts = 19, col = "grey")

```

```{r fig1disp3, echo=FALSE, out.width="50%", fig.cap="A nice image."}
spplot(world_2010, c("light_gini_lpp"), col.regions = colorRampPalette(brewer.pal(9, 'YlOrRd'))(20), 
       cuts = 19, col = "grey")
```

# Replicar Figura 2
```{r fig2disp1, echo=TRUE, out.width="50%", fig.cap="Pie de foto", warning=TRUE}
ggplot(data = swiid_df, aes(income_cat, light_gini_lpp)) + 
  geom_boxplot(colour = 'darkblue', fill = 'lightblue', size = 1) +
  labs(x = 'Income Group', y = 'Light Gini') +
  theme_minimal(base_size = 20) + xlim(c('High Income', 'Low Income', 'Middle Income'))
```


```{r fig2disp2, echo=FALSE, out.width="50%", fig.cap="Pie de foto", warning=FALSE}
ggplot(data = swiid_df, aes(income_cat, gini_disp)) + 
  geom_boxplot(colour = 'darkblue', fill = 'lightblue', size = 1) +
  labs(x = 'Income Group', y = 'Income Gini') +
  theme_minimal(base_size = 20) + xlim(c('High Income', 'Low Income', 'Middle Income'))
```

```{r fig2disp3, echo=FALSE, out.width="50%", fig.cap="Pie de foto", warning=FALSE}
ggplot(data = swiid_df, aes(region, light_gini_lpp)) + 
  geom_boxplot(colour = 'red4', fill = 'indianred1', size = 1) +
  labs(x = 'Regions', y = 'Light Gini') +
  theme_minimal(base_size = 20) 

```

```{r fig2disp4, echo=FALSE, out.width="50%", fig.cap="Pie de foto", warning=FALSE}

ggplot(data = swiid_df, aes(region, gini_disp)) + 
  geom_boxplot(colour = 'red4', fill = 'indianred1', size = 1) +
  labs(x = 'Regions', y = 'Income Gini') +
  theme_minimal(base_size = 20) 


```


# Replicar Figura 3

```{r, echo=FALSE, out.width="60%", warning=FALSE}
spplot(US_states_10, c("Estimate..Gini.Index"), par.settings=list(fontsize=list(text=20)), 
       col.regions = colorRampPalette(brewer.pal(9, 'Reds'))(50), cuts = 49, col = "transparent")

spplot(US_states_10, c("light_gini"), par.settings=list(fontsize=list(text=20)),
       col.regions = colorRampPalette(brewer.pal(9, 'Reds'))(50), cuts = 49, col = "transparent")

# La figura de EEUU por condado no se puede replicar
# spplot(US_county, c("Estimate..Gini.Index"), par.settings=list(fontsize=list(text=20)),
#        col.regions = colorRampPalette(brewer.pal(9, 'Reds'))(50), cuts = 49, col = "transparent")

cor.test(~ light_gini + Gini, df_US_states[year != 2005])

ggplot(df_US_states[year != 2005], aes(light_gini, Gini)) + geom_point() + 
  geom_smooth(method = 'lm', se = F) + labs(x = 'Light Gini', y = 'Income Gini')


```

